### 2019-11-08 initial/notes.tex
I should have known this lol

We first write out the full forms of the EOM without tidal friction. These are
taken from Kassandra's Equations 1--3 except I replace subscript $\star$ with
subscript $s$ since we are interested in the case where the spin of planet $1$
evolves with its coupling to its orbital angular momentum and perturber. We
obtain (maybe?)
\begin{align}
    \rd{\hat{s}}{t}
        &= \omega_{s1}\p*{\hat{s} \cdot \hat{l}_1}\p*{\hat{s} \times \hat{l}_1}
            - \omega_{1p}\cos(I)\p*{\hat{s} \times \hat{l}_p},\\
    \omega_{s1} &= \frac{3k_q}{2k} \p*{\frac{R_1}{a_1}}^3 s,\\
    \omega_{1p} &= \frac{3m_p}{4m_1}\p*{\frac{a_1}{a_p\p*{1 - e_p^2}}} \Omega_1.
\end{align}
Note here that $s$ is the spin frequency and $\Omega_1 =
\sqrt{GM_1/a_1^3}$ is the Keplerian orbital frequency.

@@

>>>>

We first write out the full forms of the EOM without tidal friction. These are
taken from Kassandra's Equations 1--3 except I replace subscript $\star$ with
subscript $s$ since we are interested in the case where the spin of planet $1$
evolves with its coupling to its orbital angular momentum and perturber. We
obtain (maybe?)
\begin{align}
    \rd{\hat{s}}{t}
        &= \omega_{s1}\p*{\hat{s} \cdot \hat{l}_1}\p*{\hat{s} \times \hat{l}_1}
            - \omega_{1p}\cos(I)\p*{\hat{s} \times \hat{l}_p},\\
    \omega_{s1} &= \frac{3k_q}{2k}\frac{M_*}{m_1}\p*{\frac{R_1}{a_1}}^3 s,\\
    \omega_{1p} &= \frac{3m_p}{4m_*}\p*{\frac{a_1}{a_p\sqrt{1 - e_p^2}}}^3 \Omega_1.
\end{align}
Note here that $s$ is the spin frequency and $\Omega_1 = \sqrt{GM_1/a_1^3}$ is
the Keplerian orbital frequency. We may verify that these scalings are correct:
\begin{itemize}
    \item $\omega_{1p}$ comes from the perturber being a ring that generates a
        gravitational field that is out-of-plane from the host star. We estimate
        the scalings by comparing their relative gravitational strengths. For
        this, we need that the gravitational potential of a ring inside the ring
        scales $\Phi'(r) \sim \frac{Gm_p}{a_p}\p*{\frac{r}{a_p}}^2$, and so
        \begin{equation}
            \omega_{1p} \sim \frac{\Phi'}{\Phi_0}n
                \sim \frac{m_p}{m}\p*{\frac{a_1}{a_p}}^3n.
        \end{equation}

        Recall the idea here is that we can compute some potential $\Phi(r)$ due
        to the disk $m_d$, which generally takes on form $\Phi(r, \theta) =
        -\frac{Gm_d}{a_d}\sum\limits_l a_l(\theta) \frac{r}{a_d}^l$ power
        series, and the $a_l$ are generally Legendre polynomials/spherical
        harmonics. Then, if we want the energy of an inner ring in this $\Phi$,
        we get energy $U = \int \Phi(r) m_1(r) \;\mathrm{d}r$, and if $m_1$ only
        has $l$-multipole moments then $U$ only has to be computed against
        limited terms of the expansion of $\Phi$. The way we get the expansion
        of $\Phi$ is just by expanding $\frac{1}{\abs*{\vec{r} - \vec{r}_d}}
        \sim \frac{1}{\p*{r^2 - rr_d\cos\theta + r_d^2}^{1/2}}$ as a Taylor
        series for small $\vec{r}_d$.

        I think that the vertical component of the inner ring, out of the plane
        of the disk, is not as important as the squeezing once we project it
        into the plane of the disk, but that's conjecture. It is clear that the
        quadrupole expansion of the disk potential is what drives precession,
        which gives us our scaling.

    \item $\omega_{s1}$ comes from the action of the central star on the
        quadrupole moment of the planet. This comes from applying the
        quadrupolar torque to the spin angular momentum of the planet, or
        \begin{align}
            \omega_{s1} &\sim \frac{GM_* J_2 m_1 R_1^2/a^3}{k_p m_1 R_1^2s},\\
                &\sim \frac{k_{qp}}{k_p}\frac{M_*}{m_1} \p*{\frac{R_1}{a}}^3.
        \end{align}
        Note that $J_2 = \frac{k_{qp}s^2}{Gm_1/R_1^3}$, some constant times the
        spin frequency relative to breakup. This captures the scalings.

        The argument for why this precession frequency scales like
        $\omega_{s1}\cos \theta$ is also then clear: we have some quadrupolar
        distortion of the planet, in the potential generated by the host star
        going around in a circle, so the planet projected into the plane of the
        ecliptic looks like a triaxial ellipsoid (rather than axisymmetric), and
        so will have a quadrupole moment and experience a torque that induces
        precession.
\end{itemize}

@@

\subsection{Physical Parameters}

Note that for Jupiter, $k_p \sim 0.27$ while $J_2 \sim \scinot{1.4}{-2}$ giving
$k_{qp} \sim 0.0467$ since $k_{qp} \frac{\Omega}{\sqrt{GM_p/R_p^3}} = J_2$.
Consider again the precession frequencies
\begin{align}
    \omega_{s1} &= \frac{3k_{qp}}{2k_p}\p*{\frac{R_1}{a_1}}^3s,\\
        &\approx 0.011\;\mathrm{yr^{-1}}
            \p*{\frac{5.5 k_{qp}}{k_p}}\p*{\frac{R_1}{2R_J}}^3
                \p*{\frac{a_1}{0.05\;\mathrm{AU}}}^{-3}
                \p*{\frac{s}{2\pi / 10\;\mathrm{hr}}},\\
    \omega_{1p} &= \frac{3m_p}{4m_1} \frac{a_1}{a_p\p*{1 - e_p^2}} \Omega_1,\\
        &\approx 0.017\;\mathrm{yr^{-1}}
            \frac{m_p/m_1}{0.003}
            \p*{\frac{a_1/a_p}{0.01}}
            \frac{\Omega_1}{2\pi / 3\;\mathrm{days}}
            \frac{1}{1 - e_p^2}.
\end{align}
So with a bit of tweaking, we can barely get $\omega_{1p} \lesssim \omega_{s1}$.
Recall that $\eta = \frac{\omega_{1p}\cos I}{\omega_{s1}}$, which here $\sim 1$.
We simulated above for initial conditions $\frac{s_c}{s} = \eta \sim 0.1$, which
grows as $s$ spins down. This implies $\frac{s_c}{\Omega_1} \sim 1000$
though\dots

Let's also evaluate
\begin{align}
    \rd{\theta}{t} &\approx -\frac{1}{t_a}\frac{L}{2S}\sin \theta\p*{1 -
        \frac{s}{2\Omega_1}\cos \theta},\\
        &\approx 14\;\mathrm{Myr}\p*{\frac{P}{3\;\mathrm{d}}}^{16/3}
            \p*{\frac{Q_p'}{10^7}}
            \p*{\frac{M_*}{10^3M_p}}
            \p*{\frac{a/r}{100}}^2
            \p*{\frac{s}{2\pi / 10\;\mathrm{h}}}^{-1}
            \p*{\frac{k_p}{0.25}}^{-1}
            \sin \theta\p*{1 - \frac{s}{2\Omega_1}\cos \theta}.
\end{align}
Note that typically, when trapped in CS2 and $\eta \ll 1$, $\sin \theta \approx
1$, while when at CS1, $\sin \theta \approx \sin I$. For reference, $\sin
5^\circ \approx 0.087$.

Again though, we see that if $s > \frac{2\Omega_1}{\cos\theta}$, then
$\rd{\theta}{t}$ drives towards CS2.
